Methods paper for erosion

### 2017-09-07 docs/methods/erosion_methods.tex
added QOI code for computing velocity, pressure, vorticity at target points

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
Here we focus on solving the incompressible Stokes equations with
Dirichlet boundary conditions
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  We start by defining the double-layer potential, which is the
convolution of the stresslet with a function $\eeta:\bd\Omega
\rightarrow \RR^2$,
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal at $\yy$.  For any density function $\eeta$, the double-layer
potential satisfies the incompressible Stokes equations, but the
double-layer potential can not capture velocity fields corresponding to
rigid body motions~\cite{pow-mir1987}.  A standard method to complete
the double-layer potential is to introduce a Stokeslet and rotlet
centered inside each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point inside the $k^{th}$ body,  $\rr = \xx - \cc_k$,
and $\rr^\perp = (r_2,-r_1)$.  Then, the solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ is
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\gamma_k} \eeta(\yy) \, ds_\yy, 
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align*}
  \int_{\bd\Omega} \ff \cdot \nn \, ds_\yy = 0.
\end{align*}
The result is a rank-one null space of~\eqref{eqn:completed_BIE} which
is removed by adding the rank-one modification 
\begin{align}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\nn(\yy) \, ds_\yy, \quad \xx \in \Gamma.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Pressure, and Drag} 
\label{sec:qois}
There are several quantities besides the velocity that need to be
comnputed.  First, the governing equations~\eqref{eqn:erosionModel}
require the shear stress along $\gamma_k$.  In addition, to study
characteristics of the flow, the pressure and drag must also be
computed.  Once equation~\eqref{eqn:completed_DLP} is solved for the
density function, Stokeslets, and rotlets, all of these quantities can
be computed by evaluating layer potentials that we now derive.

\subsubsection{Shear Stress}
\todo[inline]{check all signs in this section.  Fortran code assumes an
inward pointing normal}
The shear stress is defined by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ is the unit tangent vector (Figure~\ref{fig:schematic}).  We
note that for the Stokes equations, because of the no-slip boundary
condition on the bodies, the shear stress is identical to the surface
vorticity $\tau = u_y - v_x$.  However, we use
equation~\eqref{eqn:shearStress} to compute the shear stress since the
jump conditions resulting from the double-layer potential are
known~\cite{qua-bir2014a}.  The velocity field of the completed
double-layer potential~\eqref{eqn:completed_DLP} involves three terms:
the double-layer potential $\DD$, the Stokeslets $S$, and the rotlets
$R$.  To compute the shear stress, we first compute the deformation
tensor $\ssigma = \frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of each of these
terms individually.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary Integral Equation Formulation} 
\label{sec:bies}
Here we focus on solving the incompressible Stokes equations with
Dirichlet boundary conditions
\begin{alignat*}{3}
  \mu\Delta \uu &= \grad p, \qquad && \xx \in \Omega, \\
  \grad \cdot \uu &= 0,   && \xx \in \Omega, \\
  \uu &= \ff,  && \xx \in \bd\Omega.
\end{alignat*}
The viscosity $\mu$ sets the time scale, and we will assume that
$\mu=1$.  We start by defining the double-layer potential, which is the
convolution of the stresslet with a function $\eeta:\bd\Omega
\rightarrow \RR^2$,
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2} 
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
    \label{eqn:velocityDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the unit outward
normal at $\yy$.  For any density function $\eeta$, the double-layer
potential satisfies the incompressible Stokes equations, but the
double-layer potential can not capture velocity fields corresponding to
rigid body motions~\cite{pow-mir1987}.  A standard method to complete
the double-layer potential is to introduce a Stokeslet and rotlet
centered inside each interior body
\begin{align}
  S[\llambda_k,\cc_k] = \frac{1}{4\pi} \left(-\log \rho \II + 
    \frac{\rr \otimes \rr}{\rho^2} \right)\llambda_k
  \quad \text{and} \quad
  R[\xi_k,\cc_k] = \frac{\rr^\perp}{\rho^2}\xi_k,
  \label{eqn:stokeslet_rotlet}
\end{align}
where $\cc_k$ is a point inside the $k^{th}$ body,  $\rr = \xx - \cc_k$,
and $\rr^\perp = (r_2,-r_1)$.  Then, any solution of the Stokes
equations with a Dirichlet boundary condition $\ff$ can be written as
\begin{align}
  \label{eqn:completed_DLP}
  \uu(\xx) = \DD[\eeta](\xx) + \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
              \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Omega,
\end{align}
where the density function, Stokeslets, and rotlets satisfy
\begin{subequations}
\label{eqn:completed_BIE}
\begin{align}
  \ff(\xx) &= -\frac{1}{2}\eeta(\xx) + \DD[\eeta](\xx) + 
      \sum_{k=1}^{M} S[\llambda_k,\cc_k](\xx) +
      \sum_{k=1}^{M} R[\xi_k,\cc_k](\xx), \quad \xx \in \Gamma, 
      \label{eqn:DLP}\\
  \llambda_k &= \frac{1}{2\pi}\int_{\gamma_k} \eeta(\yy) \, ds_\yy, 
  \label{eqn:stokeslet} \\
  \xi_k &= \frac{1}{2\pi}\int_{\gamma_k} \yy^\perp \cdot \eeta(\yy)
  \, ds_\yy.
  \label{eqn:rotlet}
\end{align}
\end{subequations}
Finally, because of the incompressibility constraint, the boundary
condition must satisfy 
\begin{align*}
  \int_{\bd\Omega} \ff \cdot \nn \, ds_\yy = 0.
\end{align*}
The result is a rank-one null space of~\eqref{eqn:completed_BIE} which
is removed by adding the rank-one modification 
\begin{align}
  \NN_0[\eeta](\xx) = \int_{\Gamma} 
    (\nn(\xx) \otimes \nn(\yy))\nn(\yy) \, ds_\yy, \quad \xx \in \Gamma.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Computing the Shear Stress, Vorticity Pressure, and Drag} 
\label{sec:qois}
There are several quantities besides the velocity that need to be
computed.  First, the governing equations~\eqref{eqn:erosionModel}
require the shear stress along $\gamma_k$.  In addition, to study
characteristics of the flow, the vorticity, pressure, and drag must also
be computed.  Once equation~\eqref{eqn:completed_DLP} is solved for the
density function, Stokeslets, and rotlets, all of these quantities can
be computed by evaluating layer potentials that we now derive.

\subsubsection{Shear Stress}
\todo[inline]{check all signs in this section.  Fortran code assumes an
inward pointing normal}
The shear stress is defined by
\begin{align}
  \tau = \left(\nabla \uu + \nabla \uu^T \right)\nn \cdot \ss,
  \label{eqn:shearStress}
\end{align}
where $\ss$ and $\nn$ are the unit tangent and normal vectors,
respectively (Figure~\ref{fig:schematic}).  We note that for the Stokes
equations, because of the no-slip boundary condition on the bodies, the
shear stress is identical to the surface vorticity $\tau = u_y - v_x$.
However, we use equation~\eqref{eqn:shearStress} to compute the shear
stress since the jump conditions resulting from the double-layer
potential are known~\cite{qua-bir2014a}.  The velocity field of the
completed double-layer potential~\eqref{eqn:completed_DLP} involves
three terms: the double-layer potential $\DD$, the Stokeslets $S$, and
the rotlets $R$.  To compute the shear stress, we first compute the
deformation tensor $\ssigma = \frac{1}{2}(\nabla\uu + \nabla\uu^T)$ of
each of these terms individually.
